---
title: "Tutorial (basic): Modularisation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial (basic): Modularisation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  out.width = "100%"
)

library(psychTestR)
```

## About 

Modularisation is a key feature of psychTestR.
By parcelling your code into reusable components,
you can make your implementations quicker to write,
easier to read,
easier to maintain,
and easier to use.
This tutorial discusses a few strategies for modularisation in psychTestR.

## Functions

The function is the basic unit of modularisation in computer programming.
A function takes a set of inputs, performs some operation upon them,
and (optionally) returns an output:

``` {r}
f <- function(person, fruit) {
  sprintf("Greetings, %s. Would you like some %s?",
          person, fruit)
}

f(person = "fellow researcher",
  fruit = "apples")
```

When creating psychTestR tests, functions are useful for creating 
series of test elements that all share a common schema.
For example, suppose we wanted to ask someone about their
preferences for different kinds of fruit.
We could write our code like this:

``` r
join(
  NAFC_page(
    "apples", 
    "How much do you like apples, on a scale from 1 (not at all) to 7 (lots?)",
    choices = as.character(1:7), 
    arrange_vertically = FALSE
  ),
  NAFC_page(
    "pears", 
    "How much do you like pears, on a scale from 1 (not at all) to 7 (lots?)",
    choices = as.character(1:7), 
    arrange_vertically = FALSE
  ),
  NAFC_page(
    "plums", 
    "How much do you like plums, on a scale from 1 (not at all) to 7 (lots?)",
    choices = as.character(1:7), 
    arrange_vertically = FALSE
  ),
  NAFC_page(
    "apricots", 
    "How much do you like apricots, on a scale from 1 (not at all) to 7 (lots?)",
    choices = as.character(1:7), 
    arrange_vertically = FALSE
  ),
  NAFC_page(
    "pineapples", 
    "How much do you like pineapples, on a scale from 1 (not at all) to 7 (lots?)",
    choices = as.character(1:7), 
    arrange_vertically = FALSE
  ),
  NAFC_page(
    "cherries", 
    "How much do you like cherries, on a scale from 1 (not at all) to 7 (lots?)",
    choices = as.character(1:7), 
    arrange_vertically = FALSE
  ),
)

```

Clearly this is very inefficient: almost all the code is copy-pasted 
from page to page. This code is slow to read because it takes up 
so much space, and slow to maintain, because any change to the general
schema needs to be repeated six times.

Functions provide a much better way to do this. 
Consider the following:

``` r
ask_liking <- function(fruit) {
  NAFC_page(
    fruit, 
    sprintf("How much do you like %s, on a scale from 1 (not at all) to 7 (lots?)",
            fruit),
    choices = as.character(1:7), 
    arrange_vertically = FALSE
  )
}
fruits <- c("apples", "pears", "plums", "apricots", "pineapples", "cherries")
lapply(fruits, ask_liking)
```

There are a couple of new functions here that you may or may not 
have seen before:

- `sprintf` - splices variables into a template string.
- `lapply` - takes a function and applies it to each element of a list.

Our new function `ask_liking` takes one input, `fruit`, and splices it into a 
standardised schema for eliciting a 7-point liking response from the
participant.
We apply it to each element of the vector `fruits` in turn,
using the function `lapply`, 
and it returns the same list of elements that we had before,
but with many fewer lines of code.

Underneath everything, many psychTestR functions operate under 
similar principles: they programmaticaly generate test elements
according to certain schema.
If designed correctly, these functions can generalise well to many 
different applications.

## Modules

In psychTestR, 
modules are ways of wrapping sequences of test elements
into coherent logical units.^[
You may have already heard of the [modules](https://cran.r-project.org/web/packages/modules/)
package in R; despite the shared name, psychTestR modules are something different.
]
Putting a sequence of test elements into a module has three main consequences:

1. **Readability**: It makes it clear to the reader that this sequence
of test elements forms a single logical unit.
2. **Results organisation**: 
Any results generated in this module will be assigned to a special
section in the psychTestR results object,
labelled with the name of the module.
3. **Protected local environment**: 
The module will receive a fresh local environment where it can create
its own local variables (see `?set_local`).
This local environment is protected from other modules,
which is useful to avoid unexpected side effects when 
multiple modules are chained together.

Here is a simple example of module use
using the `ask_liking` function from above:

```r
ask_liking <- function(object) {
  NAFC_page(
    object, 
    sprintf("How much do you like %s, on a scale from 1 (not at all) to 7 (lots?)",
            object),
    choices = as.character(1:7), 
    arrange_vertically = FALSE
  )
}

fruits <-  c("apples", "pears", "plums")
animals <- c("dogs", "cats", "sheep")

fruit_module <- module(
  label = "fruits",
  lapply(fruits, ask_liking)
)

animal_module <- module(
  label = "animals",
  lapply(animals, ask_liking)
)

timeline <- join(
  fruit_module,
  animal_module,
  elt_save_results_to_disk(complete = TRUE),
  final_page("End.")
)

make_test(timeline)
```

If you run this code on your local computer, 
and complete the resulting test, 
then you should find your response data stored as an RDS file
in `output/results`.
Call `as.list(readRDS(file))[1:2]` in your R console,
replacing `file` with the path to the RDS file, and you should
get something like this:

```{r, echo = FALSE}
as.list(readRDS("sample-results/ex-3.rds"))[1:2]
```

As you can see, the results from the two modules have been
organised into two separate sections: one labelled `fruits`,
and one labelled `animals`.

It is often useful to wrap modules in functions that provide
a degree of customisation.
For example, suppose we wish to make `fruit_module` to have customisable length.
We could do something like this:

```{r}
fruit_module <- function(num_items) {
  stopifnot(num_items > 0, num_items <= 6)
  all_fruits <- c("apples", "pears", "plums", "apricots", "pineapples", "cherries")
  chosen_fruits <- all_fruits[1:num_items]
  lapply(chosen_fruits, ask_liking)
}
```

Now we can pass the desired number of items to our `fruit_module` function:

``` r
fruit_module(3) # a module with 3 fruits
```

Modules are not so important for one-off test implementations.
However, they do become very useful when constructing batteries
of multiple test implementations.

## Packages

A simple way of distributing R code is by sharing R source files.
The new user can read this source file into their R session
and take advantage of its contents:
for example, this source file could provide helper functions
for creating particular page types,
or it could define modules implementing entire psychological tests.

Packages are formalised ways of sharing R source files.
They stipulate particular ways of organising and documenting 
these source files, and they encapsulate these source files
within a namespace that prevents the resulting R objects
from overwriting pre-existing objects in the user's global namespace.
It is outside the scope of this tutorial to provide an introduction
to package creation in R,
but there are excellent tutorials available online,
for example [Hadley Wickham's 'R packages' book site](http://r-pkgs.had.co.nz/).
Packages are well-suited to distributing psychTestR helper functions
and modules that implement specific psychological paradigms or measures.
Here are some examples of such packages:

- [psychTestRCAT](https://pmcharrison.github.io/psychTestRCAT/) - 
helper functions for creating adaptive tests.
- [mdt](https://github.com/pmcharrison/mdt) - 
an adaptive melodic discrimination test.
- [cabat](https://github.com/pmcharrison/cabat) - 
an adaptive beat perception test.
- [mpt](https://github.com/pmcharrison/mpt) -
an adaptive mistuning perception test.
- [piat](https://github.com/pmcharrison/piat) - 
the 'Pitch Imagery Arrow Task'.
