---
title: "Tutorial (intermediate): Randomising item administration"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial (intermediate): Randomising item administration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview

This tutorial addresses the common task of randomising the order 
in which items are administered to participants.
Such techniques can be useful for avoiding order effects.

In the simplest randomisation scheme, 
every participant receives the same set of items,
and the order of these items is randomly permuted for each participant.
However, many experiments require more sophisticated randomisation schemes,
for example:

- Choosing only short questions if the participant says
they have a train to catch;
- Using the participant's initials as the seed for the randomisation 
algorithm;
- Ensuring that each item is administered the same number of times
across participants.

## Simple randomisation

Suppose we wish to administer the three survey items,
each with a yes or no response:

1. Are you afraid of dogs?
2. Are you afraid of birds?
3. Are you afraid of heights?

We can code these three items as a *tibble*:^[
A *tibble* is a nicely formatted version of a data frame, 
provide by the *tibble* package and exported by the *tidyverse* package.
]

```{r, message = FALSE, warning = FALSE}
library(psychTestR)
library(purrr)
library(tibble)

items <- tribble(
  ~ id,       ~ prompt,
  "dogs",    "Are you afraid of dogs?",
  "birds",   "Are you afraid of birds?",
  "heights", "Are you afraid of heights?"
)
items
```

There are three items in total, corresponding to the three rows of `items`.
First, let's imagine that we want each participant to receive these 
three items in a random order,
with this random order freshly determined for each participant.
We can achieve this using the psychTestR function `randomise_at_run_time`.

First, we construct a list of test elements corresponding to our 
three questions:

```{r, echo = TRUE}
item_pages <- pmap(items, function(id, prompt) {
  NAFC_page(
    label = id,
    prompt = prompt, 
    choices = c("Yes", "No")
  )
})
```

Note that we used the `pmap` function from the `purrr` package
to iterate over the rows of `items`, mapping each row to a
list element.
We then simply pass this list of pages to `randomise_at_run_time`:

``` r
timeline <- join(
  randomise_at_run_time(label = "item_order",
                        logic = item_pages),
  elt_save_results_to_disk(complete = TRUE),
  final_page("You completed the test.")
)
make_test(timeline)
```

This code should launch the test, complete with randomisation.^[
In RStudio, this means running the code with the 'Run' button,
not the 'Source' button. You can use 'Source' but then you have to wrap
the code in `shiny::runApp`.]
Results will be saved in the `output/results` directory.
Login to the admin panel using the password `demo`, 
download the results as a csv file, and
observe how the order of test elements varies between test runs,
and how the order for a given participant is saved as the variable
`results.item_order`.

## Complex randomisation

Arbitrary randomisation schemes can be implemented using the 
`order_at_run_time` function. 
For example, suppose that we wish to randomly choose between 
two item orders, 1-2-3 or 3-2-1. We can implement this as follows:

``` r
randomised <- order_at_run_time(
  label = "item_order",
  logic = item_pages,
  get_order <- function(...) {
    if (sample(2, 1) == 1) 1:3 else 3:1
  }
)

timeline <- join(
  randomised,
  elt_save_results_to_disk(complete = TRUE),
  final_page("You completed the test.")
)
make_test(timeline)
```
