---
title: "Tutorial (intermediate): Automated testing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial (intermediate): Automated testing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  out.width = "100%"
)

library(psychTestR)
```

## Introduction

Automated testing is essential in modern software development.
In the context of psychological experiment development,
automated tests help you to ensure the robustness of your experiment's implementation
without having to invest tedious hours in manual testing.

To construct an automated test in psychTestR,
you first need a test implementation saved in a file named `app.R`.
Suppose we've implemented the following minimal example:

```{r, results = "hide", message = FALSE}
library(psychTestR)

display_number <- function(i) {
  one_button_page(paste("Page", i))
}

timeline <- join(
  lapply(1:5, display_number),
  final_page("End.")
)

test <- make_test(timeline, opt = test_options("Minimal test", "password"))
```

Now make a new file, called `test.R` (or whatever you want).
Within this file, we construct an automated tester using the `AppTester` class:

```{r}
library(psychTestR)

dir <- system.file("demos/minimal-test", package = "psychTestR", mustWork = TRUE)

app <- AppTester$new(dir)

```

We've put a sample `app.R` file in the psychTestR package, 
an  the above example finds that file using `system.file`;
to point the code to your own `app.R` file, simply replace 
the `system.file` call with the path to the directory containing your
`app.R` file.

This `AppTester` object instantiates a headless browser which 
then navigates your psychTestR test, just like a real participant.
This object has various methods that are useful for interacting
with the psychTestR test and probing its internal state.
For example, we can probe the current UI text:

```{r}
app$get_ui_text()
```

We can click the "Next" button:

```{r}
app$click_next()
app$get_ui_text()
```

We can pull local variables:

```{r}
app$get_locals()
```

We can run assertions that throw an error
if a given condition is not satisfied:

```{r}
app$expect_ui_text("Page 2 Next")
```

And we can combine all of this together into an
automated script that checks that our test runs as expected:

```{r}
library(testthat) # for the expect_equal function

app <- AppTester$new(dir)

app$expect_title("Minimal test") # check the test's title

expect_equal(
  app$get_locals(),
  list(.module = NULL,
       .results_label = "results")
)

for (i in 1:5) {
  app$expect_ui_text(sprintf("Page %i Next", i))
  app$click_next()
}

app$expect_ui_text("End.")

app$stop() # shuts down the testing process
```

See `?AppTester` for more information on automated testing utilities.

## A real-world example

The above example was rather simple, but automated testing
comes into its own with more complex implementations.
Here is an example from the test suite in the psychTestR package,
designed to check that nested modules work as intended.

### The app tester

``` r
context("test_modules")

test_that("main", {
  app <- AppTester$new("apps/modules")

  app$expect_ui_text("We begin in the global environment. Next")
  app$get_locals() %>% expect_equal(list(.module = NULL,
                                         .results_label = "results"))

  app$click_next()
  app$expect_ui_text("We've now entered the parent environment. Next")
  app$get_locals() %>% expect_equal(list(.module = "parent",
                                         .results_label = "parent"))

  app$click_next()
  app$expect_ui_text("In the parent environment, we define a local variable, x = 42. Next")
  app$get_locals()$x %>% expect_equal(42)

  app$click_next()
  app$expect_ui_text("Now we enter the child environment. Next")
  app$get_locals() %>% expect_equal(list(.module = "child",
                                         .results_label = "parent.child"))

  app$click_next()
  app$expect_ui_text("We can't see x any more: x is now NULL. Next")
  app$get_locals()$x %>% expect_equal(NULL)

  app$click_next()
  app$expect_ui_text("We can set it to a new value, though: x = 65. Next")
  app$get_locals()$x %>% expect_equal(65)

  app$click_next()
  app$expect_ui_text("Now we return to the parent environment. Next")
  app$get_locals() %>% expect_equal(list(.module = "parent",
                                         .results_label = "parent",
                                         x = 42))

  app$click_next()
  app$expect_ui_text("We see that x = 42 again. Next")

  app$click_next()
  app$expect_ui_text("Now we return to the global environment. Next")
  app$get_locals() %>% expect_equal(list(.module = NULL,
                                         .results_label = "results"))

  app$click_next()
  app$expect_ui_text("We see that x is is NULL again.")
  app$get_locals()$x %>% expect_equal(NULL)

  app$get_results() %>% as.list() %>% expect_equal(list(
    parent = list(x = 42),
    parent.child = list(x = 65)
  ))

  app$stop()
})
```

### The app code

``` r
library(psychTestR)

elts <- join(
  one_button_page("We begin in the global environment."),
  module(
    "parent",

    one_button_page("We've now entered the parent environment."),
    code_block(function(state, ...) {
      x <- 42
      set_local("x", x, state)
      save_result(state, "x", x)
    }),
    one_button_page("In the parent environment, we define a local variable, x = 42."),

    module(
      "child",
      one_button_page("Now we enter the child environment."),
      one_button_page("We can't see x any more: x is now NULL."),
      code_block(function(state, ...) {
        x <- 65
        set_local("x", x, state)
        save_result(state, "x", x)
      }),
      one_button_page("We can set it to a new value, though: x = 65.")
    ),

    one_button_page("Now we return to the parent environment."),
    one_button_page("We see that x = 42 again.")
  ),

  one_button_page("Now we return to the global environment."),
  final_page("We see that x is is NULL again.")
)

make_test(elts)

```
