---
title: "Tutorial (intermediate): Editing dictionaries"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial (intermediate): Editing dictionaries}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## About

Imagine that you're using an open-source psychTestR test 
that has been created by another researcher,
and you spot a typo in one of the test items.
If this typo is hard-coded into the test implementation,
then you'll have to edit their code to fix it.
However, if the typo is instead part of their internationalisation dictionary
(see [`i18n_dict`](https://pmcharrison.github.io/psychTestR/reference/i18n_dict.html)),
then you can fix the typo without editing their code,
instead adding some code of your own when you call the test.
This tutorial demonstrates this functionality.

## Instructions

Let's begin by making a simple test with a typo.

```{r, include = FALSE, echo = FALSE}
library(psychTestR, quietly = TRUE)
```

```{r}
# Load the psychTestR package
library(psychTestR)

# Define a dictionary with a typo
my_dict <- i18n_dict$new(data.frame(key = "key1",
                                    en = "Welcoom to the test!"))

# The test itself is distributed as a function,
# which might itself be distributed as a package
my_test <- function(dict = my_dict) {
  make_test(new_timeline(
    list(final_page(i18n("key1"))),
    dict = dict
  ))
}
```

If we run this test, we see the typo.

``` r
my_test()
```

To fix this problem, we begin by inspecting the dictionary object
to find out which component we need to change.
In real-life applications, 
this dictionary object should be exported from the package
distributing your test, 
and you should be able to find it by inspecting the documentation.
We can inspect the dictionary using the `as.data.frame` method.

```{r}
my_dict$as.data.frame()
```

Inspecting this dictionary shows us that we want to edit the entry
with key `key1` and language code `en`.
We can do this as follows:

```{r}
my_dict$edit(key = "key1",
             language = "en",
             new = "Welcome to the test!")
```

Unlike most R functions, this `edit` function is *destructive*:
it updates the original object, even without an assignment operator.
So, if we inspect our `my_dict` object, we see that it has now changed:

```{r}
my_dict$as.data.frame()
```

## Real-life example

To make this more concrete, here's a real-life example.
Let's install the melody discrimination test of
[Harrison & Müllensiefen (2017)](https://github.com/pmcharrison/mdt):

``` r
if (!requireNamespace("devtools")) install.packages("devtools")
devtools::install_github("pmcharrison/mdt")
```

We can demo the test as follows:

``` r
mdt::demo_mdt()
```

Now, suppose I want to replace the word 'Next' with the word 'Proceed'.
Inspecting the documentation, I see that the dictionary is stored 
in the object `mdt::mdt_dict`.
I inspect this dictionary:

```r
View(mdt::mdt_dict$as.data.frame())
```

and see, after some searching, that the key for the 'Next' button 
is `AMDI_0016_I_0001_1`.

(Note: If you get the error "attempt to apply non-function", 
you probably have an outdated version of `psychTestR` or `mdt` - 
reinstalling these packages should fix the problem.)

I then edit the dictionary as follows:

``` r
mdt::mdt_dict$edit(key = "AMDI_0016_I_0001_1", 
                   language = "en",
                   new = "Proceed")
```

I can check this worked by asking the dictionary to translate the key:

``` r
mdt::mdt_dict$translate(key = "AMDI_0016_I_0001_1", 
                        language = "en")
```

And now, if we run the test, we should see our changes:

``` r
mdt::demo_mdt()
```

Note that this edit only persists for the length of your R session:
if you restart, it'll revert to the original.

## Feedback

Have you got feedback about this documentation? 
Please submit it to the [issues tracker](https://github.com/pmcharrison/psychTestR/issues).
